---
title: "ENGAGE_LME_Task_Activation"
Author: Xue Zhang
date: "04/13/2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
toc: yes
toc_float: yes
code_folding: hide
---

### load useful packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width= 2, fig.asp = 0.8, fig.pos = '!h')
# saved <- options(repr.plot.width=3, repr.plot.height=2) #  ,repr.plot.asp = 0.8
pacman::p_load(pacman, rio, lme4, lmerTest,tidyverse, sjPlot, coefplot2, performance, see, broom.mixed, kableExtra, janitor, ggeffects, dplyr, gridExtra, qqplotr, emmeans, pbkrtest, knitr,ggpubr, here, table1, psych)
options(kableExtra.auto_format = TRUE)
# options(knitr.table.format = 'markdown')
# load functions used by this text written by me
source(here("R/ggplot_the_model.R"))

# use here from the here package
here <- here::here
```

### define subfunctions
create table via kable
```{r warning=FALSE, paged.print=TRUE, comment= NA, message=FALSE, results='asis'}
## create table via kable
create_table <- function(data_table, caption_var, footnote_var){
  data_table <- head(as.data.frame(data_table), Inf) %>%
    mutate(across(is.numeric, ~ round(., 2)))
  if ("Pr(>F)" %in% colnames(data_table)){
     data_table <- rename(data_table, p.value = `Pr(>F)`) %>%
     mutate(p.value = case_when(
    p.value < 0.001 ~ paste0(as.character(p.value), "***"),
    TRUE ~ paste0(p.value)
  )) %>%
    mutate(p.value = case_when(
    p.value < 0.01 ~ paste0(as.character(p.value), "**"),
    TRUE ~ paste0(p.value)
  )) %>%
    mutate(p.value = case_when(
    p.value < 0.05 ~ paste0(as.character(p.value), "*"),
    TRUE ~ paste0(p.value)
  )) %>%
    mutate(p.value = case_when(
    p.value < 0.10 ~ paste0(as.character(p.value), "."),
    TRUE ~ paste0(p.value)
  ))
  }
  
  kable_plot <- data_table %>%
    kable()
  #   # rename(p.value = `Pr(>F)`) %>%
  #   kbl(caption = caption_var, format = "html") %>%
  #   kable_classic(lightable_options = "hover",full_width = F, html_font = "Arial",  font_size = 10) %>%
  # footnote(general = footnote_var,general_title = "---") #%>%
  #   # save_kable(file = paste(save_name, sep = ""), self_contained = T)
  cat('\n')
  print(kable_plot)
  cat('\n')
  }

```
run_ggplot_the_effects
```{r fig.keep = "all", fig.align = "center"} 
# to be noted the fig setting here doesn;t work, change the options in ggplot_the_effects
## ggplot the effects
run_ggplot_the_effects <- function(emtrends_results, title_var){
  # calculate confidence interval
  margin_confint <- confint(emtrends_results$contrasts, level=0.95)
  emtrends_contrast_sum <- summary(emtrends_results$contrasts)
  emtrends_contrast_sum$lower.CL <- margin_confint$lower.CL
  emtrends_contrast_sum$upper.CL <- margin_confint$upper.CL
  
  marginal_plot <- ggplot_the_effects(
    fit = lme_model,
    fit_pairs = emtrends_contrast_sum,
    effect_label = title_var)
  options(repr.plot.width=0.5, repr.plot.height= 2) # ,repr.plot.asp = 0.8 doesn't work
  print(marginal_plot + theme(axis.text = element_text(size = 8)))
}
```


### read data

be careful of setting the option of contrast for LME modeling

```{r}
# read fmri data
Datadir = "/Users/xuezhang/Documents/Stanford/Projects/ENGAGE/Analysis/engage/SPM/Level2/MixedModel/111_fMRI_stats_spikesonly_FD_fromFile_GO_NO_GO/NogovsGo/SPSI/MixedModel_SPSI/DataforR"
data_csv <- import(paste(Datadir,"/Data4Model2_3_MixedModel_SPSI.csv",sep = ""))

# read medication prescription information
med_data <- import('/Users/xuezhang/Documents/Stanford/Projects/ENGAGE/Analysis/engage/CSV/medication_data.csv')

med_data_long <- med_data %>%
  mutate('24MO' = med_12mo_binary) %>%
  rename('6MO' = 'med_6mo_binary', '12MO' = 'med_12mo_binary', subjects = Subjects) %>%
  select(subjects, '6MO', '12MO', '24MO') %>%
  pivot_longer(-c(subjects), names_to = "Time", values_to = 'med_binary') %>%
  mutate(Time = factor(Time, levels = c('6MO','12MO','24MO'))) %>%
  drop_na() 

# combine medication prescription data with fmri data
data_csv <- merge(data_csv, med_data_long, by = c('subjects', 'Time'),all = TRUE) 

data_csv$Time <- factor(data_csv$Time, levels = c('6MO','12MO','24MO'))
data_csv$Group <- factor(data_csv$Group, levels = c('U-care','I-care'))
data_csv$gender <- factor(data_csv$gender, levels = c('Female','Male'))
data_csv$subjects <- factor(data_csv$subjects)
options(contrasts = c("contr.sum","contr.poly"))
head(data_csv,1)
```

### ROI list

```{r}
ROI_names <- c('lDLPFC2','rIPL_1','Precuneus2','lFusiform','rIPL_2','lIPL')
```

### Identifying cognitive control circuit as a neural mechanisam engaged by the I-CARE intervention

```{r warning=FALSE, paged.print=TRUE, comment= NA, message=FALSE, fig.keep = "all", results='asis'}
# footnote for plot
footnote_var_anova <- as.character("Signif. codes:  0 `'***'` 0.001 `'**'` 0.01 `'*'` 0.05 `'.'` 0.1 `' '` 1")

footnote_var_emtrend <- as.character("Results are averaged over the levels of: Group, gender 
Degrees-of-freedom method: kenward-roger 
P value adjustment: tukey method for comparing a family of 3 estimates")
for (i_name in 1:4) {
  #lme modeling
  ROI_name_i <- ROI_names[i_name]
  #fmri variable
  fmri <- paste(ROI_name_i,"_change",sep = "")
  #formula
  formula <-  as.formula(paste("SPSI_change ~ SPSI_baseline + BMI_baseline + ", fmri, " + ",fmri,":Time + ",fmri,":Group + ",fmri, ":Time:Group + age + gender + (1|subjects)", sep = ""))
  # # formula that includes medication prescription as a covariate
  # formula <-  as.formula(paste("SPSI_change ~ SPSI_baseline + BMI_baseline + ", fmri, " + ",fmri,":Time + ",fmri,":Group + ",fmri, ":Time:Group + med_binary + age + gender + (1|subjects)", sep = ""))
  # # formula that includes behavioral performance as covariates
  # formula <-  as.formula(paste("SPSI_change ~ SPSI_baseline + BMI_baseline + ", fmri, " + ",fmri,":Time + ",fmri,":Group + ",fmri, ":Time:Group + fmri_fFA_change + fmri_fFT_change + age + gender + (1|subjects)", sep = ""))
  #lmer modeling
  lme_model <- lmer(formula, data = data_csv, REML = FALSE)
  #ROI name
  print(ROI_name_i)
  cat("\n\n\n\n")
  #check model
  # print(check_model(lme_model))
  cat("\n\n\n\n")
  #summary model
  print("Linear Mixed Model")
  print(lme_model)
  cat("\n\n\n\n")
  #show anova -- goes to supplement, needs modification
  anova_result <- anova(lme_model,type = 'marginal')
  create_table(anova_result, "Marginal Analysis of Variance Table with Satterthwaite's method", footnote_var_anova)
  cat("\n\n\n\n")

    #plot SPSI ~ fmri x group, only lines
    formula_plot_2 = as.formula(paste("Group ~ ", fmri, sep = ""));
    a <-  emmip(lme_model,  formula_plot_2 , cov.reduce = range, CIs = T) + labs(title = ROI_name_i, x = "\u0394 NoGo > Go fMRI activation", y = "\u0394 SPSI") + theme_classic() + theme(legend.position = "top") #ylim(c(-2.5,2)) + xlim(c(-1.5,1.5)) +
    print(a + scale_colour_brewer(palette = "Set1") + scale_fill_brewer(palette = "Set1"))
    cat("\n\n\n\n")
    #plot marginal effect
    print("Marginal Effects")
    emtrends_results <- emtrends(lme_model, pairwise ~ Group, var = fmri)
    create_table(emtrends_results, "Post-hoc Marginal Estimates", footnote_var_emtrend)
    run_ggplot_the_effects(emtrends_results, "")
  
}
```


## Identifying cognitive control circuit engagement at 2 months as a predictor of later behavioral improvement at 6, 12, and 24 months
```{r warning=FALSE, paged.print=TRUE, comment= NA, message=FALSE, fig.keep = "all", results='asis'}
## r fig.asp=0.8, fig.width=4,
for (i_name in 5:6) {
  #lme modeling
  ROI_name_i <- ROI_names[i_name]
  #fmri variable
  fmri <- paste(ROI_name_i,"_2MO_change",sep = "")
  #formula
  formula <-  as.formula(paste("SPSI_change ~ SPSI_baseline + ", fmri, " + ",fmri,":Time + ",fmri,":Group + ",fmri, ":Time:Group + age + gender + (1|subjects)", sep = ""))
  
  # # formula that includes medication prescription as a covariate
  # formula <-  as.formula(paste("SPSI_change ~ SPSI_baseline + BMI_baseline + ", fmri, " + ",fmri,":Time + ",fmri,":Group + ",fmri, ":Time:Group + med_binary + age + gender + (1|subjects)", sep = ""))
  # # formula that includes behavioral performance as covariates
  # formula <-  as.formula(paste("SPSI_change ~ SPSI_baseline + BMI_baseline + ", fmri, " + ",fmri,":Time + ",fmri,":Group + ",fmri, ":Time:Group + fmri_fFA_change + fmri_fFT_change + age + gender + (1|subjects)", sep = ""))
  #lmer modeling
  lme_model <- lmer(formula, data = data_csv, REML = FALSE)
  #ROI name
  print(ROI_name_i)
  cat("\n\n\n\n")
  #check model
  # print(check_model(lme_model))
  cat("\n\n\n\n")
  #summary model
  print("Linear Mixed Model")
  print(summary(lme_model))
  cat("\n\n\n\n")
  #show anova -- goes to supplement, needs modification
  anova_result <- anova(lme_model,type = 'marginal')
  cat("\n\n\n\n")
  create_table(anova_result, "Marginal Analysis of Variance Table with Satterthwaite's method", footnote_var_anova)
  #general indicator
  emtrends_results <- emtrends(lme_model, pairwise ~ Time, var = fmri)
  create_table(emtrends_results, "Post-hoc Marginal Estimates", footnote_var_emtrend)
  run_ggplot_the_effects(emtrends_results, "Pairwise Comparison")
  #plot SPSI ~ fmri x time, with scatters
  formula_plot_1 = as.formula(paste("Time ~ ", fmri, sep = ""));
  print(emmip(lme_model,  formula_plot_1 , cov.reduce = range, CIs = T) +
    geom_point(aes_string(x = fmri, y = "SPSI_change", colour = "Time", fill = "Time"), data = data_csv, pch = 16, width = 0.1, alpha = 0.5) + labs(title = ROI_name_i, x = "\u0394 NoGo > Go fMRI activation", y = "\u0394 SPSI") + theme_classic() + theme(legend.position = "top")+ scale_colour_brewer(palette = "Set1") + scale_fill_brewer(palette = "Set1")) #ylim(c(-2.5,2)) + xlim(c(-1.5,1.5)) +
  cat("\n\n\n\n")

}
```


### fmri GoNoGo performance vs. SPSI
```{r warning=FALSE, paged.print=TRUE, comment= NA, message=FALSE, fig.keep = "all", results='asis'}
variables <- c('fmri_fFT_change', 'fmri_fFA_change')
for (var in variables) {
  print(var)
  cat("\n")
 formula <- as.formula(paste("SPSI_change ~ SPSI_baseline + BMI_baseline + ", var, " + ", var, ":Time + ", var, ":Group + ", var, ":Time:Group + age + gender + (1|subjects)", sep = ""))
  lme_model <- lmer(formula, data = data_csv, REML = FALSE)
  anova_result <- anova(lme_model,type = 'marginal')    
  print(anova_result)
}

```

```{r}
# Clear figures
#dev.off()

# Clear environment
rm(list = ls()) 

# Clear packages
p_unload(all)  # Remove all add-ons

# Clear console
cat("\014")  # ctrl+L

# Clear mind :)

```




