---
title: "ENGAGE_prediction"
output: html_document
---


### load useful packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
# knitr::opts_chunk$set(fig.width= 3, fig.asp = 0.8, fig.pos = '!h') 
# saved <- options(repr.plot.width=3, repr.plot.height=2) #  ,repr.plot.asp = 0.8
pacman::p_load(pacman, rio, lme4, lmerTest,tidyverse, sjPlot, coefplot2, performance, see, broom.mixed, kableExtra, janitor, ggeffects, dplyr, gridExtra, qqplotr, emmeans, pbkrtest, knitr,ggpubr, here, table1, psych, cvms, groupdata2, caret, pROC,glmnet, caret)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = 'markdown')

# use here from the here package
here <- here::here

# load functions used by this text written by me
source(here("R/ggplot_the_model.R"))


```
### define subfunctions
create table via kable
```{r warning=FALSE, paged.print=TRUE, comment= NA, message=FALSE, results='asis'}
## create table via kable
create_table <- function(data_table, caption_var, footnote_var){
  data_table <- head(as.data.frame(data_table), Inf) %>%
    mutate(across(is.numeric, ~ round(., 2)))
  if ("Pr(>F)" %in% colnames(data_table)){
     data_table <- rename(data_table, p.value = `Pr(>F)`) %>%
     mutate(p.value = case_when(
    p.value < 0.001 ~ paste0(as.character(p.value), "***"),
    TRUE ~ paste0(p.value)
  )) %>%
    mutate(p.value = case_when(
    p.value < 0.01 ~ paste0(as.character(p.value), "**"),
    TRUE ~ paste0(p.value)
  )) %>%
    mutate(p.value = case_when(
    p.value < 0.05 ~ paste0(as.character(p.value), "*"),
    TRUE ~ paste0(p.value)
  )) %>%
    mutate(p.value = case_when(
    p.value < 0.10 ~ paste0(as.character(p.value), "."),
    TRUE ~ paste0(p.value)
  ))
  }
  
  kable_plot <- data_table %>%
    # rename(p.value = `Pr(>F)`) %>%
    kbl(caption = caption_var, format = "html") %>%
    kable_classic(full_width = F, html_font = "Arial",  font_size = 10) %>%
  footnote(general = footnote_var,general_title = "---") #%>%
    # save_kable(file = paste(save_name, sep = ""), self_contained = T)
  cat('\n')
  print(kable_plot)
  cat('\n')
  }

```
run_ggplot_the_effects
```{r fig.keep = "all", fig.align = "center"} 
# to be noted the fig setting here doesn;t work, change the options in ggplot_the_effects
## ggplot the effects
run_ggplot_the_effects <- function(emtrends_results, title_var){
  # calculate confidence interval
  margin_confint <- confint(emtrends_results$contrasts, level=0.95)
  emtrends_contrast_sum <- summary(emtrends_results$contrasts)
  emtrends_contrast_sum$lower.CL <- margin_confint$lower.CL
  emtrends_contrast_sum$upper.CL <- margin_confint$upper.CL
  
  marginal_plot <- ggplot_the_effects(
    fit = glme_modal,
    fit_pairs = emtrends_contrast_sum,
    effect_label = title_var)
  options(repr.plot.width=0.5, repr.plot.height= 2) # ,repr.plot.asp = 0.8 doesn't work
  print(marginal_plot + theme(axis.text = element_text(size = 8)))
  cat('\n')
}

```

### read data

be careful of setting the option of contrast for LME modeling

```{r}
# read fmri data
Datadir = "/Users/xuezhang/Documents/Stanford/Projects/ENGAGE/Analysis/engage/SPM/Level2/MixedModel/111_fMRI_stats_spikesonly_FD_fromFile_GO_NO_GO/NogovsGo/SCL_20/MixedModel_SCL20/DataforR"
Figdir = "/Users/xuezhang/Documents/Stanford/Projects/ENGAGE/Analysis/SPM/Figures/Symptom"
data_csv <- import(paste(Datadir,"/Data4Model2_3_MixedModel_SCL20.csv",sep = ""))
data_csv$Time <- factor(data_csv$Time, levels = c('6MO','12MO','24MO'))
data_csv$Group <- factor(data_csv$Group, levels = c('U-care','I-care'))
data_csv$gender <- factor(data_csv$gender, levels = c('Female','Male'))
data_csv$subjects <- factor(data_csv$subjects)
options(contrasts = c("contr.sum","contr.poly"))
head(data_csv,1)
```
### ROI list

```{r}
ROI_names <- c('rIPL3','lSPL')
```

## continous outcome prediction
```{r warning=FALSE, paged.print=TRUE, comment= NA, message=FALSE, fig.keep = "all", results='asis'}
# r fig.asp=0.8, fig.width=4, 
for (i_name in 1:2) {

  #lme modeling
  ROI_name_i <- ROI_names[i_name]
  #fmri variable
  fmri <- paste(ROI_name_i,"_2MO_change",sep = "")
    # filter data
  data_filter <- data_csv %>%
    select(fmri, subjects, gender, age, Group, Time, SCL_20_change, SCL_20_baseline) %>%
    drop_na()
  
  # cross validation
  # Set seed for reproducibility
  for (i in 10) {
    set.seed(i)
    cat("\n")
    print(paste("This is Seed ",i,sep = ""))
  # select data for modeling
  data_filter <- data_csv %>%
    select(lSPL_2MO_change, rIPL3_2MO_change, subjects, gender, age, Group, Time, SCL_20_change, SCL_20_baseline) %>%
    drop_na()  
  # Fold data
  k_fold <- 5
  data_fold <- fold(
    data = data_filter, k = k_fold,
    cat_col = 'Group',id_col = 'subjects') %>% #
    arrange(.folds)
 # cross validation
  CV1 <- cross_validate(
  data = data_fold,
  formulas = c("SCL_20_change ~ age + gender + (1|subjects)","SCL_20_change ~ SCL_20_baseline + age + gender + (1|subjects)", paste("SCL_20_change ~ SCL_20_baseline + ", fmri, " + age + gender + (1|subjects)", sep = "")),
  fold_cols = '.folds',
  family = 'gaussian',
  REML = FALSE
)

  print(CV1 %>% select_metrics() %>% mutate(across(is.numeric, ~ round(., 2))) %>% kbl() %>% kable_classic(full_width = F, html_font = "Arial",font_size = 13))
# # Nested results from the different folds
print(CV1$Results[[1]] %>% mutate(across(is.numeric, ~ round(., 2)))%>% kbl() %>% kable_classic(full_width = F, html_font = "Arial",font_size = 13))
print(CV1$Results[[2]] %>% mutate(across(is.numeric, ~ round(., 2)))%>% kbl() %>% kable_classic(full_width = F, html_font = "Arial",font_size = 13))
print(CV1$Results[[3]] %>% mutate(across(is.numeric, ~ round(., 2)))%>% kbl() %>% kable_classic(full_width = F, html_font = "Arial",font_size = 13)) 

  r_test <- 0
  r_test_baseline <- 0
  r_test_control <- 0
  # cv_performance <- data.frame()
  formula_control <- "SCL_20_change ~ age + gender + (1|subjects)"
  formula_baseline <- "SCL_20_change ~ SCL_20_baseline + age + gender + (1|subjects)"
  formula_sim <- paste("SCL_20_change ~ SCL_20_baseline + ", fmri, " + age + gender + (1|subjects)", sep = "")
  for (i_fold in 1:k_fold) {
    print(paste("This is fold ",i_fold,sep = ""))
    data_ifold <- data_fold %>%
      filter(.folds != i_fold)
    # print(data_ifold %>% head(Inf) %>% kable())
    data_test <- data_fold %>%
      filter(.folds == i_fold)
    lme_model_ifold_control <- lmer(formula_control, data = data_ifold, REML = FALSE)
    lme_model_ifold_baseline <- lmer(formula_baseline, data = data_ifold, REML = FALSE)
    lme_model_ifold <- lmer(formula_sim, data = data_ifold, REML = FALSE)
    
    
    anova_result <- anova(lme_model_ifold,type = 'marginal')
    cat("\n\n\n\n")
    # create_table(anova_result, "Marginal Analysis of Variance Table with Satterthwaite's method", footnote_var_anova)
    lme_predict <- predict(lme_model_ifold, newdata = data_test, allow.new.levels = TRUE)
    lme_predict_baseline <- predict(lme_model_ifold_baseline, newdata = data_test, allow.new.levels = TRUE)
    lme_predict_control <- predict(lme_model_ifold_control, newdata = data_test, allow.new.levels = TRUE)
    lme_rmse <- sqrt(mean((lme_predict - data_test$SCL_20_change)^2))
    lme_rmse_baseline <- sqrt(mean((lme_predict_baseline - data_test$SCL_20_change)^2))
    lme_rmse_control <- sqrt(mean((lme_predict_control - data_test$SCL_20_change)^2))
    print(paste("RMSE = ", lme_rmse, sep = ""))
    cat("\n")
    print(paste("RMSE Baseline = ", lme_rmse_baseline, sep = ""))
    cat("\n")
    print(paste("RMSE Control = ", lme_rmse_control, sep = ""))
    cat("\n")
    r_test <- r_test + cor(lme_predict, data_test$SCL_20_change,  method = "pearson", use = "complete.obs")
    r_test_baseline <- r_test_baseline + cor(lme_predict_baseline, data_test$SCL_20_change,  method = "pearson", use = "complete.obs")
    r_test_control <- r_test_control + cor(lme_predict_control, data_test$SCL_20_change,  method = "pearson", use = "complete.obs")
    knit_print(cor.test(lme_predict, data_test$SCL_20_change,  method = "pearson", use = "complete.obs"))
    cat("\n")
    knit_print(cor.test(lme_predict_baseline, data_test$SCL_20_change,  method = "pearson", use = "complete.obs"))
    cat("\n")
    knit_print(cor.test(lme_predict_control, data_test$SCL_20_change,  method = "pearson", use = "complete.obs"))
    cat("\n")
    

  }
  print(paste("mean r = ", round(r_test/k_fold, 3), sep = ""))
  print(paste("mean baseline r = ", round(r_test_baseline/k_fold, 3), sep = ""))
  print(paste("mean control r = ", round(r_test_control/k_fold, 3), sep = ""))
  
  }
  
}
```


## Predicting binary outcome: response
```{r warning=FALSE, paged.print=TRUE, comment= NA, message=FALSE, fig.keep = "all", results='asis'}

## Logistic regression only for 6month
set.seed(5)
  y_all_test <- matrix(0, nrow = nrow(data_fold), ncol = 1) # combine all test subjects
y_all_test_con <- matrix(0, nrow = nrow(data_fold), ncol = 1) # combine all test subjects
y_all_test_baseline_con <- matrix(0, nrow = nrow(data_fold), ncol = 1) # combine all test subjects
y_all_test_baseline <- matrix(0, nrow = nrow(data_fold), ncol = 1) # combine all test subjects
performance_test <- matrix(0, nrow = 12, ncol = k_fold)
performance_test_baseline <- matrix(0, nrow = 12, ncol = k_fold)
performance_test_control <- matrix(0, nrow = 12, ncol = k_fold)
set.seed(2)
  # select data for modeling
  data_filter <- data_csv %>%
    select(lSPL_2MO_change, rIPL3_2MO_change, subjects, gender, age, Group, Time, Response, Remission, SCL_20_baseline) %>%
    mutate(Response= 2 - as.numeric(Response), Remission= 2- as.numeric(Remission)) %>%
    drop_na()  #%>% filter(Time == "6MO") #, fmri
  data_filter_6MO <- data_filter %>%
    filter(Time == "6MO")
  # Fold data
  k_fold <- 2
  data_fold <- fold(
    data = data_filter, k = k_fold,
    cat_col = 'Group',id_col = 'subjects') %>% #
    arrange(.folds)
  
  
  
  for (i_fold in 1:k_fold) { #
    print(paste("This is fold ",i_fold,sep = ""))
    data_ifold <- data_fold %>%
      filter(.folds != i_fold, Time =="6MO")
    # print(data_ifold %>% head(Inf) %>% kable())
    data_test <- data_fold %>%
      filter(.folds == i_fold)
    
    x_train <- data_ifold%>%
  select(lSPL_2MO_change,rIPL3_2MO_change, gender,age, SCL_20_baseline)%>%
  data.matrix()%>%
    .[,-1]
  
  # baseline symptoms
  x0_train <- data_ifold%>%
  select(age,gender, SCL_20_baseline)%>%
  data.matrix()%>%
    .[,-1]
    
  # y
  y_train <- data_ifold%>%
  select(Response)%>%
    data.matrix()%>%
    .[,-1]
  
  ## testing data
  x_test <- data_test%>%
  select(lSPL_2MO_change,rIPL3_2MO_change, gender,age, SCL_20_baseline)%>%
  data.matrix()%>%
    .[,-1]
  # baseline symptoms
  x0_test <- data_test%>%
  select(age,gender, SCL_20_baseline)%>%
  data.matrix()%>%
    .[,-1]
  # y
  y_test <- data_test%>%
  select(Response)%>%
  data.matrix()%>%
    .[,-1]

  
   fit_logistic <- glm(formula[3], data = data_ifold, family = "binomial")
   y_train_pred_s <- predict(fit_logistic, newdata = data_ifold, type='response') %>% # 1se
    data.matrix()
  y_test_pred_s <- predict(fit_logistic, newdata = data_test, type='response') %>% # 1se
    data.matrix()
  
    my_roc <- roc(y_train, y_train_pred_s) #roc(y_test, y_test_pred_s)
  my_th <- coords(my_roc, "best", ret = "threshold") %>%
    data.matrix()
    
    ind_1 <- y_test_pred_s>=my_th[1]
  ind_0 <- y_test_pred_s<my_th[1]
  y_test_pred_s_c <- y_test_pred_s
  y_test_pred_s_c[ind_1]=1
  y_test_pred_s_c[ind_0]=0
  y_all_test[data_fold$.folds == i_fold] <- y_test_pred_s_c # combine all sub
  y_all_test_con[data_fold$.folds == i_fold] <- y_test_pred_s
  print(con_mat <- confusionMatrix(as.factor(y_test_pred_s_c),as.factor(y_test)))
  cat("\n")
  plot(fit_logistic)
  performance_test[, i_fold] = c(con_mat$overall[1],con_mat$byClass)
  
  roc(y_test, y_test_pred_s, percent=F,   boot.n=1000, ci.alpha=0.9, stratified=FALSE, plot=TRUE, grid=TRUE, show.thres=TRUE, legacy.axes = TRUE, reuse.auc = TRUE,
# print.thres = c(0.30,0.35, 0.40, 0.45,0.48, 0.50,0.55, 0.60),#
print.auc = TRUE, print.thres.col = "blue", ci=TRUE, ci.type="bars", print.thres.cex = 0.7 ) # , main = paste("ROC curve using","(N = 78")    ",length(y_test),")
  
# prediction using symptoms
  
 fit_logistic_baseline<- glm(formula[2], data = data_ifold, family = "binomial")

  y_train_pred_baseline_s = predict(fit_logistic_baseline, newdata = data_ifold, type='response') %>% # 1se
    data.matrix()
  y_test_pred_baseline_s = predict(fit_logistic_baseline, newdata = data_test, type='response') %>% # 1se
    data.matrix()
    
  my_roc <- roc(y_train, y_train_pred_baseline_s)#roc(y_test, y_test_pred_baseline_s)
  my_th <- coords(my_roc, "best", ret = "threshold")%>%
    data.matrix()
  
  
  ind_1 <- y_test_pred_baseline_s>=my_th[1]
  ind_0 <- y_test_pred_baseline_s<my_th[1]
  
  y_test_pred_baseline_s_c <- y_test_pred_baseline_s
  y_test_pred_baseline_s_c[ind_1]=1
  y_test_pred_baseline_s_c[ind_0]=0
   y_all_test_baseline[data_fold$.folds == i_fold] <- y_test_pred_baseline_s_c
  y_all_test_baseline_con[data_fold$.folds == i_fold] <- y_test_pred_baseline_s
  print(con_mat_baseline <- confusionMatrix(as.factor(y_test_pred_baseline_s_c),as.factor(y_test)))
  plot(fit_logistic_baseline)
  cat("\n")
  performance_test_baseline[, i_fold] = c(con_mat_baseline$overall[1],con_mat_baseline$byClass)
  
  roc(y_test, y_test_pred_baseline_s, percent=F,   boot.n=1000, ci.alpha=0.9, stratified=FALSE, plot=TRUE, grid=TRUE, show.thres=TRUE, legacy.axes = TRUE, reuse.auc = TRUE,
# print.thres = c(0.30,0.35, 0.40, 0.45,0.48, 0.50,0.55, 0.60),#
print.auc = TRUE, print.thres.col = "blue", ci=TRUE, ci.type="bars", print.thres.cex = 0.7 ) # , main = paste("ROC curve using","(N = 78)")   ",length(y_test),"
  
  
}
for (i in c(1:k_fold)) {
  print(paste("This is  ",i,sep = ""))
  print(mean(performance_test[i,]))
print(mean(performance_test_baseline[i,]))
}



# performance for combined all subjects
# with fmri
print(con_mat_all_fmri <- confusionMatrix(as.factor(y_all_test),as.factor(data_fold$Response)))
print(con_mat_all_fmri$overall %>%
        kable())
performance_test_all_fmri <-  c(con_mat_all_fmri$overall[1],con_mat_all_fmri$byClass)

# only baseline
print(con_mat_all_baseline <- confusionMatrix(as.factor(y_all_test_baseline),as.factor(data_fold$Response)))
  cat("\n")
  performance_test_all_baseline <-  c(con_mat_all_baseline$overall[1],con_mat_all_baseline$byClass)

# ROC
# tiff(file=paste(Figdir, "/ROC_logistic_regression.tiff", sep = ""), width=3, height=3, res=300, units="in")
par(pty = "s")
roc(data_fold$Response, y_all_test_con, plot=TRUE, ci.alpha=0.95,  ci=FALSE, ci.type="bars", legacy.axes = TRUE, percent = TRUE, xlab= "False Positive Percentage", ylab = "True Positive Percentage", col="#a50f15", lwd=4, print.auc=TRUE, print.auc.x=80, print.auc.y=20)
plot.roc(data_fold$Response,y_all_test_baseline_con, ci.alpha=0.95,  ci=FALSE, ci.type="bars", percent = TRUE, col="#252525", lwd=4, print.auc=TRUE, add=TRUE, print.auc.x=80, print.auc.y=10)
par(pty = "m")
# dev.off()
```


```{r}
# Clear figures
#dev.off()

# Clear environment
rm(list = ls()) 

# Clear packages
p_unload(all)  # Remove all add-ons

# Clear console
cat("\014")  # ctrl+L

# Clear mind :)